### 2017-02-14 docs/plan.tex
Additional notes on addressing model
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@2765 40876441-3496-0410-a613-a2fca35b2ae2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  plan.tex
%
%  Software Configuration Management Plan
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  Kevin Stratford (kevin@epcc.ed.ac.uk)
%  (c) 2011-2015 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The \textit{Ludwig} code is designed to study the hydrodynamic properties
of simple and complex fluids based around the numerical solution of the
Navier-Stokes equations. Complex fluids are dealt with via a free-energy
based finite difference approach, which couples explicitly to the
hydrodynamics.
Components of \textit{Ludwig} include the main program, unit and
regression tests, and a small number support libraries, and a small
number of utility programs used to prepare
input and post-process output. The Plan is limited to these
software components.

\textit{Ludwig} is specifically designed for parallel computers and
relies on the Message Passing Interface \cite{mpi-standard}. The
code will also run in serial, and is independent of
the platform it is run on. The Plan is not concerned with
hardware or system management activities.

@@

This section is based on the IEEE standard for Software Configuration
Management Plans IEEE 828-2012 \cite{ieee-828-2012}, and follows the format
set out therein.
Relevant references are included and can be found in the main References
section at the end of the document.

\subsection{Management}

@@

\subsubsection{Responsibilities}

While the Soft Matter Physics Group is responsible for the overall
scientific direction and development of the code and concomitant
priorities, all the Software Configuration Management activities will
be the responsibility of EPCC.

% IEEE section recheck what this actually means
% \subsubsection{Procedures}
% There are currently no external constraints on The Plan.

@@

\subsection{Activities}

\subsubsection{Identification}

The material under control of the project includes three sets of files:
this documentation, the source code, and the test suite which is used
to monitor the status of the code.
Control of project files is via a single on-line repository which
uses the subversion (SVN) revision control system. The SVN repository
is currently located at

@@

All access to the code is via the CCPForge repository. CCPForge identifies
three different levels of access:

\textit{Administrators:} allows administration of the site itself and
registered members;

\textit{Developers:} members who have write access to the source code repository;

\textit{Users:} have read only access to the source code via SVN (either
as a member or via anonymous download). Member users also have access to
the tracker.

As CCPForge is not physically under the control of EPCC, a
complete dump of the repository to a local resource is undertaken on a
regular basis for security.

@@

%\begin{thebibliography}{99}
%\bibitem{ieee-828-2012} IEEE Standard 828-2011 ``IEEE Standard for
%Software Configuration Management Plans''
%\bibitem{mpi-standard} Message Passing Interface Forum. MPI: A Message
%Passing Interface Standard Version 1.3 (1998).
%\bibitem{paraview} Paraview is a standard visualisation package. 
%\bibitem{apacheAPR} See, e.g., the Apache Portable Runtime Project ``APR's Version Numbering''  https://apr.apache.org/versioning.html (Accessed 2015).
%\end{thebibliography}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  plan.tex
%
%  Software Configuration Management Plan
%
%  See, for example, Annex D of IEEE 828-2012
%  IEEE Standard for Configuration Management in Systems and Software
%  Engineering
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  (c) 2011-2015 The University of Edinburgh
%  Contributing authors:
%  Kevin Stratford (kevin@epcc.ed.ac.uk)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The \textit{Ludwig} code is designed to study simple and complex fluids
based around the numerical solution of the Navier-Stokes equations.
Components of \textit{Ludwig} include the main program, unit and
regression tests, and a small number support libraries, and a small
number of utility programs used to prepare
input and post-process output. The Plan is limited to these
software components.

% IEEE section to be completed
\subsubsection{Relationship to organisation and other projects}

@@

This section is based on the IEEE standard for Configuration
Management IEEE 828-2012 \cite{ieee-828-2012}, and follows the format
set out therein.
Relevant references are included and can be found in the main References
section at the end of the document.
%

% Criteria for identication

\subsection{Identification}

The material under control of the project includes a number of sets
of files: this documentation, the source code, the build and test
suite which is used to monitor the status of the code, and so on.
Control of project files is via a single on-line repository which
uses the subversion (SVN) revision control system. The SVN repository
is currently located at

Specifically, configuration items are taken to include files in the
trunk of the repository

\texttt{http://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk}

Other branches in the repository are not considered to be 
under configuration management.

% IEEE section
% Limitations and assumptions
% Deliberately empty at the moment.

\subsection{Responsibility and authority}

While the Soft Matter Physics Group is responsible for the overall
scientific direction and development of the code and concomitant
priorities, all Configuration Management activities will
be the responsibility of EPCC at The University of Edinburgh.

\subsection{Project organisation}

@@

@@

@@

%All access to the code is via the CCPForge repository. CCPForge identifies
%three different levels of access:

%\textit{Administrators:} allows administration of the site itself and
%registered members;

%\textit{Developers:} members who have write access to the source code repository;

%\textit{Users:} have read only access to the source code via SVN (either
%as a member or via anonymous download). Member users also have access to
%the tracker.

%As CCPForge is not physically under the control of EPCC, a
%complete dump of the repository to a local resource is undertaken on a
%regular basis for security.

@@

%\begin{thebibliography}{99}

%Software Configuration Management Plans''
%\bibitem{mpi-standard} Message Passing Interface Forum. MPI: A Message
%Passing Interface Standard Version 1.3 (1998).
%\bibitem{paraview} Paraview is a standard visualisation package.

### 2017-02-14 targetDP/doc/data_parallel_execution.tex

\chapter{Data Parallel Execution}

@@

The \verb+__targetILP__+ syntax is used, within a \verb+__targetTLP__+
region, to specify that the proceeding block of code should be
executed in parallel and mapped to instruction level parallelism
(ILP), where the extent of the ILP is defined by the virtual vector length (VVL) in the targetDP implementation. 
\begin{comment}
need to document VVL in more detail somewhere and link.
\end{comment}

@@

\subsection{Example}
**TO DO**
\subsection{Implementation}
\subsubsection{C}
**TO DO**
\subsubsection{CUDA}
**TO DO**

>>>>

\chapter{Data Parallel Execution}\label{sec:dpexecution}

This chapter specifices the data parallel execution functionality in targetDP.

@@

The \verb+__targetILP__+ syntax is used, within a \verb+__targetTLP__+
region, to specify that the proceeding block of code should be
executed in parallel and mapped to instruction level parallelism
(ILP), where the extent of the ILP is defined by the virtual vector length (VVL) in the targetDP implementation (see \ref{sec:exmodel}).

@@

### 2017-02-14 targetDP/doc/memory_management.tex

\chapter{Memory Management}

@@

@@

@@

\begin{itemize}
\item \verb+variableName+: The name of the variable or array.
\item \verb+type+: The type of variabe or array.
\end{itemize}

\subsection{Example}
**TO DO**.
\subsection{Implementation}
\subsubsection{C}
Holds no value
\subsubsection{CUDA}
\verb+__constant__+

@@

\subsection{Example}
**TO DO**.

@@

\subsection{Example}
**TO DO**.
\subsection{Implementation}
\subsubsection{C}
**TO DO**.
\subsubsection{CUDA}
**TO DO**.

@@

\subsection{Example}
**TO DO**.

\subsection{Implementation}
\subsubsection{C}
**TO DO**.
\subsubsection{CUDA}
**TO DO**.

\newpage
\section{copyToTargetBoundary3D}

The \verb+copyToTargetBoundary3D+ function copies the data corresponding to the boundaries of a 3D lattice from the host to the target.

\subsection{Syntax}
\begin{verbatim}
void copyToTargetBoundary3D(void *targetData, const void *data, size_t extent3D[3], size_t nField, size_t offset, size_t depth);
\end{verbatim}

\begin{itemize}
\item \verb+targetData+: A pointer to the destination array on the target.
\item \verb+data+: A pointer to the source array on the host.
\item \verb+extent3D+: An array of 3 integers corresponding to the 3D dimensions of the lattice.
\item \verb+nFields+: The number of fields per lattice site.
\item \verb+offset+: The number of sites from the lattice edge at which each boundary face should start.
\item \verb+depth+: The depth of each boundary face.
\end{itemize}

\subsection{Example}
**TO DO**.
\subsection{Implementation}
\subsubsection{C}
**TO DO**.
\subsubsection{CUDA}
**TO DO**.

\newpage
\section{copyFromTargetBoundary3D}

The \verb+copyFromTargetBoundary3D+ function copies the data corresponding to the boundaries of a 3D lattice from the target to the host.

\subsection{Syntax}
\begin{verbatim}
void copyFromTargetBoundary3D(void *data, const void *targetData, size_t extent3D[3], size_t nField, size_t offset, size_t depth);
\end{verbatim}

\begin{itemize}
\item \verb+data+: A pointer to the destination array on the host.
\item \verb+targetData+: A pointer to the source array on the target.
\item \verb+extent3D+: An array of 3 integers corresponding to the 3D dimensions of the lattice.
\item \verb+nFields+: The number of fields per lattice site.
\item \verb+offset+: The number of sites from the lattice edge at which each boundary face should start.
\item \verb+depth+: The depth of each boundary face.
\end{itemize}

\subsection{Example}
**TO DO**.
\subsection{Implementation}
\subsubsection{C}
**TO DO**.
\subsubsection{CUDA}
**TO DO**.
\newpage
\section{copyToTargetPointerMap3D}

@@

\subsection{Syntax}
\begin{verbatim}
void copyToTargetPointerMap3D(void *targetData, const void *data, 
           size_t extent3D[3], size_t nField, 
           int includeNeighbours, void** pointerArray);
\end{verbatim}

\begin{itemize}
\item \verb+targetData+: A pointer to the destination array on the target.
\item \verb+data+: A pointer to the source array on the host.
\item \verb+extent3D+: An array of 3 integers corresponding to the 3D dimensions of the lattice.
\item \verb+nField+: The number of fields per lattice site.
\item \verb+includeNeighbours+: A boolean switch to specify whether each included site should also have it's neighbours included (in the 19-point 3D stencil).
\item \verb+pointerArray+: An array of \verb+nSite+ pointers, where \verb+nSite+ is the total number of lattice sites. Each lattice site should be included unless the pointer corresponding to that site is \verb+NULL+.  
\end{itemize}

\subsection{Example}
**TO DO**.
\subsection{Implementation}
\subsubsection{C}
**TO DO**.
\subsubsection{CUDA}
**TO DO**.

@@

\subsection{Syntax}
\begin{verbatim}
void copyFromTargetPointerMap3D(void *data, const void *targetData, 
           size_t extent3D[3], size_t nField, 
           int includeNeighbours, void** pointerArray);
\end{verbatim}

\begin{itemize}
\item \verb+data+: A pointer to the destination array on the host.
\item \verb+targetData+: A pointer to the source array on the target.
\item \verb+extent3D+: An array of 3 integers corresponding to the 3D dimensions of the lattice.
\item \verb+nField+: The number of fields per lattice site.
\item \verb+includeNeighbours+: A boolean switch to specify whether each included site should also have it's neighbours included (in the 19-point 3D stencil).
\item \verb+pointerArray+: An array of \verb+nSite+ pointers, where \verb+nSite+ is the total number of lattice sites. Each lattice site should be included unless the pointer corresponding to that site is \verb+NULL+.  
\end{itemize}

\subsection{Example}
**TO DO**.

\subsection{Implementation}
\subsubsection{C}
**TO DO**.
\subsubsection{CUDA}
**TO DO**.

>>>>

\chapter{Memory Management}\label{chap:memmanage}

This chapter specifices the memory management functionality in targetDP.

@@

\newpage

\section{targetMallocUnified}

The \verb+targetMallocUnified+ function allocates unified memory that can be accessed on the host or the target. This should be used with caution since it may result in poor performance relative to use of \verb+targetMalloc+.

\subsection{Syntax}
\begin{verbatim}
void targetMallocUnified(void** targetPtr, size_t n);
\end{verbatim}

\subsection{Example}
Analogous to Line \ref{example:scalelaunch:targetMalloc} in Figure \ref{fig:scalelaunch} in Section \ref{chapter:examples}.
\subsection{Implementation}
\subsubsection{C}
\verb+malloc+
\subsubsection{CUDA}
\verb+cudaMallocManaged+

\newpage
\section{targetCallocUnified}

The \verb+targetCallocUnified+ function allocates, and initialises to zero, unified memory that can be accessed on the host or the target. This should be used with caution since it may result in poor performance relative to use of \verb+targetCalloc+.

\subsection{Syntax}
\begin{verbatim}
void targetCallocUnified(void** targetPtr, size_t n);
\end{verbatim}

\subsection{Implementation}
\subsubsection{C}
\verb+calloc+
\subsubsection{CUDA}
\verb+cudaMallocManaged+ followed by \verb+cudaMemset+

@@

\newpage
\section{copyDeepDoubleArrayToTarget}

The \verb+copyDeepDoubleArrayToTarget+ function copies an array of double precision values from the host to the target, where the array is contained within another object.

\subsection{Syntax}
\begin{verbatim}
void copyDeepDoubleArrayToTarget(void* targetObjectAddress,
    void* hostObjectAddress,void* hostComponentAddress,size_t n);
\end{verbatim}

\begin{itemize}
\item \verb+targetObjectAddress+: A pointer to the target copy of the object that contains the data array.
\item \verb+hostObjectAddress+: A pointer to the host copy of the object that contains the data array.
\item \verb+hostComponentAddress+: A pointer to the host copy of the start of the array contained within the object.
\item \verb+n+: The number of elements to be copied.
\end{itemize}

\subsection{Implementation}
\subsubsection{C}
Pointer arithmetic to determine memory locations, followed by \verb+memcpy+
\subsubsection{CUDA}
Pointer arithmetic to determine memory locations, followed by \verb+cudaMemcpy+

\newpage
\section{copyDeepDoubleArrayFromTarget}

The \verb+copyDeepDoubleArrayTFromTarget+ function copies an array of double precision values from the target to the host, where the array is contained within another object.

\subsection{Syntax}
\begin{verbatim}
void copyDeepDoubleArrayFromTarget(void* hostObjectAddress,
   void* targetObjectAddress,void* hostComponentAddress,size_t n);
\end{verbatim}

\begin{itemize}
\item \verb+hostObjectAddress+: A pointer to the host copy of the object that contains the data array.
\item \verb+targetObjectAddress+: A pointer to the target copy of the object that contains the data array.
\item \verb+hostComponentAddress+: A pointer to the host copy of the start of the array contained within the object.
\item \verb+n+: The number of elements to be copied.
\end{itemize}

\subsection{Implementation}
\subsubsection{C}
Pointer arithmetic to determine memory locations, followed by \verb+memcpy+
\subsubsection{CUDA}
Pointer arithmetic to determine memory locations, followed by \verb+cudaMemcpy+

\newpage
\section{targetZero}

The \verb+targetZero+ function sets a (double precision) array on the target to zero.

\subsection{Syntax}
\begin{verbatim}
void targetZero(double* targetData, size_t n);
\end{verbatim}

\begin{itemize}
\item \verb+targetData+: A pointer to the array on the target.
\item \verb+n+: The number of elements in the array.
\end{itemize}

\subsection{Implementation}
\subsubsection{C}
A loop to zero each element.
\subsubsection{CUDA}
A kernel to zero each element.

\newpage
\section{targetSetConstant}

The \verb+targetSetConstant+ function sets each element of a (double precision) array on the target to the specified constant value.

\subsection{Syntax}
\begin{verbatim}
void targetSetConstant(double* targetData, double value, size_t n);
\end{verbatim}

\begin{itemize}
\item \verb+targetData+: A pointer to the array on the target.
\item \verb+value+: The value.
\item \verb+n+: The number of elements in the array.
\end{itemize}

\subsection{Implementation}
\subsubsection{C}
A loop to set each element.
\subsubsection{CUDA}
A kernel to set each element.

@@

\begin{itemize}
\item \verb+variableName+: The name of the variable or array.
\item \verb+type+: The type of variable or array.
\end{itemize}

\subsection{Example}
See Line \ref{example:scalekernel:targetConst} in Figure \ref{fig:scalelaunch} in Section \ref{chapter:examples}.
\subsection{Implementation}
\subsubsection{C}
Holds no value
\subsubsection{CUDA}
\verb+__constant__+

@@

@@

@@

%% \newpage
%% \section{copyToTargetBoundary3D}

%% \subsection{Description}

%% The \verb+copyToTargetBoundary3D+ function copies the data corresponding to the boundaries of a 3D lattice from the host to the target.

%% \subsection{Syntax}
%% \begin{verbatim}
%% void copyToTargetBoundary3D(void* targetData, const void* data, 
%%               size_t extent3D[3], size_t nField, 
%%               size_t offset, size_t depth);
%% \end{verbatim}

%% \begin{itemize}
%% \item \verb+targetData+: A pointer to the destination array on the target.
%% \item \verb+data+: A pointer to the source array on the host.
%% \item \verb+extent3D+: An array of 3 integers corresponding to the 3D dimensions of the lattice.
%% \item \verb+nFields+: The number of fields per lattice site.
%% \item \verb+offset+: The number of sites from the lattice edge at which each boundary face should start.
%% \item \verb+depth+: The depth of each boundary face.
%% \end{itemize}

%% \newpage
%% \section{copyFromTargetBoundary3D}

%% \subsection{Description}

%% The \verb+copyFromTargetBoundary3D+ function copies the data corresponding to the boundaries of a 3D lattice from the target to the host.

%% \subsection{Syntax}
%% \begin{verbatim}
%% void copyFromTargetBoundary3D(void* data, const void* targetData, 
%%                  size_t extent3D[3], size_t nField, 
%%                  size_t offset, size_t depth);
%% \end{verbatim}

%% \begin{itemize}
%% \item \verb+data+: A pointer to the destination array on the host.
%% \item \verb+targetData+: A pointer to the source array on the target.
%% \item \verb+extent3D+: An array of 3 integers corresponding to the 3D dimensions of the lattice.
%% \item \verb+nFields+: The number of fields per lattice site.
%% \item \verb+offset+: The number of sites from the lattice edge at which each boundary face should start.
%% \item \verb+depth+: The depth of each boundary face.
%% \end{itemize}

\newpage
\section{copyToTargetPointerMap3D}

@@

\subsection{Syntax}
\begin{verbatim}
void copyToTargetPointerMap3D(void* targetData, const void* data, 
           size_t extent3D[3], size_t nField, 
           int includeNeighbours, void** pointerArray);
\end{verbatim}

\begin{itemize}
\item \verb+targetData+: A pointer to the destination array on the target.
\item \verb+data+: A pointer to the source array on the host.
\item \verb+extent3D+: An array of 3 integers corresponding to the 3D dimensions of the lattice.
\item \verb+nField+: The number of fields per lattice site.
\item \verb+includeNeighbours+: A Boolean switch to specify whether each included site should also have it's neighbours included (in the 19-point 3D stencil).
\item \verb+pointerArray+: An array of \verb+nSite+ pointers, where \verb+nSite+ is the total number of lattice sites. Each lattice site should be included unless the pointer corresponding to that site is \verb+NULL+.  
\end{itemize}

@@

\subsection{Syntax}
\begin{verbatim}
void copyFromTargetPointerMap3D(void* data, const void* targetData, 
           size_t extent3D[3], size_t nField, 
           int includeNeighbours, void** pointerArray);
\end{verbatim}

\begin{itemize}
\item \verb+data+: A pointer to the destination array on the host.
\item \verb+targetData+: A pointer to the source array on the target.
\item \verb+extent3D+: An array of 3 integers corresponding to the 3D dimensions of the lattice.
\item \verb+nField+: The number of fields per lattice site.
\item \verb+includeNeighbours+: A Boolean switch to specify whether each included site should also have it's neighbours included (in the 19-point 3D stencil).
\item \verb+pointerArray+: An array of \verb+nSite+ pointers, where \verb+nSite+ is the total number of lattice sites. Each lattice site should be included unless the pointer corresponding to that site is \verb+NULL+.  
\end{itemize}

### 2017-02-14 targetDP/doc/memory_model.tex

\chapter{Memory model}

\section{Model overview}
The targetDP model draws a distinction between the memory spaces
accessed by the host and target, such that code executed on the host
always accesses the host memory space, and code executed on the target
(i.e. within target functions) always accesses the target memory space. The
host memory space can be initialised using regular \verb+C/C+++
functionality, and the targetDP API provides the functionality
necessary to manage the target memory space and transfer data between
the host and target. For each data-parallel data structure, the
programmer should create both host and target copies, and should
update these from each other as and when required.

@@

\subsection{Target memory model}
The team of threads performing the execution of target functions
should always act on target data structures. These can take one of 3
forms:
\begin{enumerate}
\item Those created using the targetDP memory allocation API
  functions. These are shared between all threads in the team, where
  each thread should use its unique index to access the portion of data
  for which it is responsible.
\item Those created using the targetDP constant data management
  functionality. These are read-only and normally used for relatively
  small amounts of constant data.
\item Those declared within the body of a target function. These are private
  to each thread in the team and should be used for temporary scratch structures.
\end{enumerate}

\begin{comment}
For 3, at the moment any data declared in the function but above the targetTLP keyword will be private for GPU threads but shared for CPU (since outside parallel region). So either need to make this clearer above, or somehow move OpenMP parallel region to target launch. 
\end{comment}

>>>>

\chapter{Memory model}\label{sec:memmodel}

\section{Model overview}
The targetDP model draws a distinction between the memory space
accessed by the host and that accessed by the target. Even although
there is a trend towards ``unified'' address spaces, on which this
distinction is not strictly required for applications to run successfully, such
visibility at the application level is often crucial to allow good
performance when running on those target architectures that have
associated high-bandwidth memory systems (such as GPU and Intel Xeon
Phi architectures). In the targetDP model, it is assumed that code
executed on the host always accesses the host memory space, and code
executed on the target (i.e. within target functions) always accesses
the target memory space. The host memory space can be initialised
using regular \verb+C/C+++ functionality, and the targetDP API
provides the functionality necessary to manage the target memory space
and transfer data between the host and target. For each data-parallel
data structure, the programmer should create both host and target
copies, and should update these from each other as and when required.

@@

\subsection{Target memory model}
The team of threads performing the execution of target functions
should always act on target data structures. These data structures can
take one of 3 forms:
\begin{enumerate}
\item Those created using the targetDP memory allocation API
  functions. These are shared between all threads in the team, where
  each thread should use its unique index to access the portion of data
  for which it is responsible.
\item Those created using the targetDP constant data management
  functionality. These are read-only and normally used for relatively
  small amounts of constant data.
\item Those declared within the body of a target function. These are private
  to each thread in the team and should be used for temporary scratch structures.
\end{enumerate}

%\begin{comment}
%For 3, at the moment any data declared in the function but above the targetTLP keyword will be private for GPU threads but shared for CPU (since outside parallel region). So either need to make this clearer above, or somehow move OpenMP parallel region to target launch. 
%\end{comment}
